---
output:
  pdf_document: default
  html_document: default
---

=======
---
title: "Lab1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

# Introduction
A research community studied government policies' influence in fostering a healthy broadband market. While it was revealed that U.S. is among the few countries that did not adopt open access policies out of all OECD countries, network owners insist that attempts to regulate prices or mandate greater penetration will reduce incentives to invest in modern equipment, and thus, there is a trade-off among price, speed, and penetration.
We are analyzing 3 sets of data - Price, Penetration, Speed - to examine if there are positive/negative relationship between Speed(quality of network) and price and/or penetration.
More explicitly:
1) Does the network quality(Speed) falls if price is held low? Positive relationship?
2) Does the network quality(Speed) falls if penetration is forced high? Negative relationship?

## Setup
First, we load some libraries and the data into R.  We do this using dplyr to take advantage of its table manipulation methods.  For easy access, we join the three datasets together.  A simple full_join is sufficient.


```{r load_data, warning=FALSE}
library(dplyr)
library(car)
library(ggplot2)

t1 = as_tibble(read.csv('Penetration.csv', stringsAsFactors = F))
t2 = as_tibble(read.csv('Price.csv', stringsAsFactors = F))
t3 = as_tibble(read.csv('Speed.csv', stringsAsFactors = F))

t = t1 %>% full_join(t2) %>% full_join(t3)
str(t)
```

## Cleaning the Data

We see a couple of issues with the data:
1. There are two blank rows at the bottom, which We filter out of the table.  
2. The 'X' variable seems to be empty, so we omit that column.
3. Due to a typo in the original data, Both 'Country.Code' and 'Country.code' exist.  We omit the latter.
4. There are percent signs, dollar signs, and commas being used as decimal separators.  There's probably some R function better than read.csv for cleaning these columns, but let's use some regular expressions just for fun. 

```{r clean_data, warning=FALSE}
clean_percent = function(x) as.numeric(gsub('%', '', x))
clean_dollar = function(x) as.numeric(gsub('\\$', '', x))
clean_commas = function(x) as.numeric(gsub(',', '', x))

t = t %>%
    filter(Country != '') %>%
    select(-c(X, Country.code)) %>%
    mutate(Growth.in.3G.penetration = clean_percent(Growth.in.3G.penetration)) %>%
    mutate(Percent.of.population.in.urban.areas = clean_percent(Percent.of.population.in.urban.areas))

for (c in grep('Price', names(t))) t[c] = unlist(lapply(t[c], clean_dollar))
for (c in grep('Maximum|Average|Standard|Median|X90p', names(t))) t[c] = unlist(lapply(t[c], clean_commas))
t$Country = factor(t$Country)
t$Country.Code = factor(t$Country.Code)
str(t)
```

Now we have a tidy data set.  All the columns hold either numeric or integer data with the exceptions of 'Country' and 'Country.Code'.

# Univariate Analysis of Key Variables

For looking at key variables by country, we will use ggplot since it makes better barplots than the base package.

First let's take a look at household penetration by country.
```{r}
library(ggplot2)
ggplot(t, aes(x=reorder(Country, Household.penetration..OECD), y=Household.penetration..OECD)) +
    geom_bar(stat='identity') +
    geom_text(aes(label=Household.penetration..OECD), nudge_y = 3, size = 2) +
    labs(title = 'Household Penetration by Country', y = 'Household Penetration (%)', x = 'Country') +
    coord_flip()
```
Korea has the highest Household penetration with 94.13%.  The United States is pretty middle-of-the-road with 50.8%.  Turkey has the lowest penetration at 1.73%.

Let's also take a look at average speed.

```{r}
ggplot(t, aes(x=reorder(Country, Average.actual.speed..Akamai..kbps.), y=Average.actual.speed..Akamai..kbps.)) +
    geom_bar(stat='identity') +
    geom_text(aes(label=Average.actual.speed..Akamai..kbps.), nudge_y = 3, size = 2) +
    labs(title = 'Average Speed by Country', y = 'Average Actual Speed (kbps)', x = 'Country') +
    coord_flip()
```

We immediately notice a few countries with missing data.  It is worth noting that the countries with missing data (Turkey, Slovak Republic, Poland, and Hungary) were also very low in household penetration.  I think that this may be an error of some sort.  Let's take a look at the same plot but with some filtered rows.

```{r}
ggplot(filter(t, !is.na(Average.actual.speed..Akamai..kbps.)), 
        aes(x=reorder(Country, Average.actual.speed..Akamai..kbps.), y = Average.actual.speed..Akamai..kbps.)) +
    geom_bar(stat='identity') +
    geom_text(aes(label=Average.actual.speed..Akamai..kbps.), nudge_y = 0, size = 2) +
    labs(title = 'Average Speed by Country (NAs removed)', y = 'Average Actual Speed (kbps)', x = 'Country') +
    coord_flip()
```

Here we see that Korea, which had the highest penetration, now has the highest speed at 15.239 kbps.  United States has a medium speed at 3.913 kbps. Luxemborg has lowest speed at 2.497 kbps.  

Finally, let's take a look at some prices.  We're not really sure of the currency unit, but it's all relative anyways.  We have chosen to leave NA's in the data set.  It is interesting to consider why certain countries are missing prices are certain speed tiers.  Is it a data collection error or do they simply not provide those speeds?  Note that we do not actually know where the cut off points for these speed tiers are in the original dataset (e.g. very high vs. high).
```{r}
ggplot(t, aes(x=reorder(Country, Price.for.very.high.speeds..combined), y = Price.for.very.high.speeds..combined)) +
    geom_bar(stat='identity') +
    geom_text(aes(label=Price.for.very.high.speeds..combined), nudge_y = 3, size = 2) +
    labs(title = 'Price for Very High Speeds by Country', y = 'Price for Very High Speeds', x = 'Country') +
    coord_flip()
```

```{r}
ggplot(t, aes(x=reorder(Country, Price.for.high.speeds..combined), y = Price.for.high.speeds..combined)) +
    geom_bar(stat='identity') +
    geom_text(aes(label=Price.for.high.speeds..combined), nudge_y = 5, size = 2) +
    labs(title = 'Price for High Speeds by Country', y = 'Price for High Speeds', x = 'Country') +
    coord_flip()
```

```{r}
ggplot(t, aes(x=reorder(Country, Price.for.med.speeds..combined), y = Price.for.med.speeds..combined)) +
    geom_bar(stat='identity') +
    geom_text(aes(label=Price.for.med.speeds..combined), nudge_y = 3, size = 2) +
    labs(title = 'Price for Medium Speeds by Country', y = 'Price for Medium Speeds', x = 'Country') +
    coord_flip()
```

```{r}
ggplot(t, aes(x=reorder(Country, Price.for.low.speeds..combined), y = Price.for.low.speeds..combined)) +
    geom_bar(stat='identity') +
    geom_text(aes(label=Price.for.low.speeds..combined), nudge_y = 2, size = 2) +
    labs(title = 'Price for Low Speeds by Country', y = 'Price for Low Speeds', x = 'Country') +
    coord_flip()
```
The takeaway here is that the ranking of countries in price changes quite a lot when varying speeds.

## Average Network Speed

We now examine the overall distribution of speeds, with no segregation among countries.  The plotted speeds are measured by Akamai.

```{r}
summary(t$Average.actual.speed..Akamai..kbps., na.rm = T)
hist(t$Average.actual.speed..Akamai..kbps., breaks = 20,
     main = 'Histogram of Average Speed', xlab = 'Average Speed (kbps)')
```
There seems to be an outlier at high end (it's Korea with its ridiculously good internet), which makes it difficult to examine the distribution of the values where values are clustered. Let's remove the outlier.
```{r}
t_sub1 <- subset(t, Average.actual.speed..Akamai..kbps. < 8000, na.rm = T)
summary(t_sub1$Average.actual.speed..Akamai..kbps.)
hist(t_sub1$Average.actual.speed..Akamai..kbps., breaks = 20,
     main = 'Histogram of Average Speed (Outlier Removed)', xlab = 'Average Speed (kbps)')
```
We can now notice that mean value has become close to the median, and that most countries have network speed between 1000 and 7000.  The distribution of speed actually looks fairly normal.

## Household penetration
```{r}
summary(t$Household.penetration..OECD)
hist(t$Household.penetration..OECD, breaks = 20,
     main = 'Histogram of Household Penetration', xlab = 'Household Penetration (%)')
```
We see that most of the countries' household penetration lies between 25% and 80%.

## Price for various speeds
### Price for very high speed network
```{r}
summary(t$Price.for.very.high.speeds..combined, na.rm = T)
hist(t$Price.for.very.high.speeds..combined, breaks = 20,
     main = 'Histogram of Price for Very High Speeds', xlab = 'Price for Very High Speeds')
```
This histogram shows that price for very high speeds are everywhere, without any regular dispersion.

### Price for high speed network
```{r}
summary(t$Price.for.high.speeds..combined, na.rm = T)
hist(t$Price.for.high.speeds..combined, breaks = 20,
     main = 'Histogram of Price for High Speeds', xlab = 'Price for High Speeds')
```
We can see that the outlier (Poland) above 200 makes examination of distribution difficult.  Let us take a look at the distribution sans that one data point.
```{r}
t_sub2 <- subset(t_sub1, Price.for.high.speeds..combined < 200, na.rm = T)
hist(t_sub2$Price.for.high.speeds..combined, breaks = 20,
     main = 'Histogram of Price for High Speeds (Outlier Removed)', xlab = 'Price for High Speeds')
```
From this revised subset, we can see that most countries charge 30 to 70 dollars for high speed internet varying diversely.

### Price for Med speed network
```{r}
summary(t$Price.for.med.speeds..combined)
hist(t$Price.for.med.speeds..combined, breaks = 20,
     main = 'Histogram of Price for Medium Speeds', xlab = 'Price for Medium Speeds')
```
Compared to high speed networks, medium speed network shows more clustered around 30-50 dollars.

### Price for Low speed network
```{r}
summary(t$Price.for.low.speeds..combined)
hist(t$Price.for.low.speeds..combined, breaks = 20,
     main = 'Histogram of Price for Low Speeds', xlab = 'Price for Low Speeds')
```
Low price network also has a similarly shaped histogram to the medium speed one, as opposed to high speed one.

#Analysis of Key Relationships

##Scatterplot Matrix

To examine relationships among the key variables, we first draw some scatterplot matrices.
```{r}
scatterplotMatrix( ~t$Average.actual.speed..Akamai..kbps. + t$Household.penetration..OECD, 
                   diagonal = 'hist',
                   main = 'Average Speed vs Household Penetration')
```
We see a slightly positive relationship between penetration and speed.  Countries with higher household penetration also have higher speeds.  The outlier to the far right does seem to be changing the slope of the relationship a bit.

Next, let's draw a scatterplot matrix for speed against the various price points.

```{r}
scatterplotMatrix( ~ t_sub2$Household.penetration..OECD + 
                       t_sub2$Price.for.very.high.speeds..combined + 
                       t_sub2$Price.for.high.speeds..combined + 
                       t_sub2$Price.for.med.speeds..combined + 
                       t_sub2$Price.for.low.speeds..combined, 
                   diagonal = 'hist',
                   main = 'Average Speed vs Prices')
```

Perhaps uninterestingly, the prices at various speed breakpoints track each other very well.  A country with cheap internet at one breakpoint has cheap internet at the others; one with expensive internet has expensive internet at all breakpoints as well.  The more interesting relationship is that speed seems to be negatively related to all four tiers of pricing.  That is, the faster the internet a country has, the lower the pricing.  

We can clearly see:
1. Positive relationship between Speed and Penetration;
2. Negative relationships between Speed and Prices, more so with Price for med and low speed networks than Price for high speed networks.
The reason why there is weaker relationship between Speed and price for high/veryhigh than Speed and price for low/med is because prices for high/veryhigh network speeds are dispersed more broadly than low/med network.
This is completely the opposite to the argument of the network owners.


# Analysis of secondary effects
## Create the regression line to see the relationship.

From the previous scatter plot matrix we discovered:
1. There is a positive relationship between the household penetration rate and the actual speed, only when the consumers are paying for high or very high speed. 
2. House penetration rate and the medium and low price displayed negative relationship. The result suggested that the lower the price consumer paid, the less house penetration rate the signal is.

To further investigate the relationship between price and speed, we Created the regression line.

```{r}
plot(t$Household.penetration..OECD, factor = 2, t$Price.for.very.high.speeds..combined, factor = 2,
     xlab = "House Penetration", ylab = "Very High Speed Price",
     main = "House Penetration and Price of Very High Speed")
abline(lm(t$Household.penetration..OECD ~ t$Price.for.very.high.speeds..combined))

plot(t$Household.penetration..OECD, factor = 2, t$Price.for.high.speeds..combined, factor = 2,
     xlab = "House Penetration", ylab = "High Speed Price",
     main = "House Penetration and Price of High Speed")
abline(lm(t$Household.penetration..OECD ~ t$Price.for.high.speeds..combined))

plot(t$Household.penetration..OECD, factor = 2, t$Percent.of.population.in.urban.areas, factor = 2,
     xlab = "House Penetration", ylab = "Population % in Urban Areas",
     main = "House Penetration and Population in Urban Areas")

plot(t$Average.actual.speed..Akamai..kbps., factor = 2, t$Percent.of.population.in.urban.areas, factor = 2,
     xlab = "Average Speed", ylab = "Population % in Urban Areas",
     main = "Average Speed and Population in Urban Areas")
```

From the graph, we see there are relationship between average actual speed and the percenta of population in urban areas and the price.
To further investigate the relationship we use linear regression to see if the average actual speed change according to these two variables.

```{r}
lm(t$Household.penetration..OECD ~ t$Percent.of.population.in.urban.areas)
```

On average, the household penetration rate is 29.6. For every percent increase in the population in urban areas, the household penetration rate increase by 0.21.

Next, we examine the average actual speed.

```{r}
lm(t$Average.actual.speed..Akamai..kbps. ~ t$Percent.of.population.in.urban.areas)
```

On average, the average actual speed is 2451.79. For every percent increase in the population, the average speed increase by 22.81. This is somehow questionable because we expect the more user in the area, the lower the average speed is. The result showed the opposite.

Then we look at the actual average speed with the price for very high speed internet, the high speed,the medium, and the low speed.

```{r}
lm(t$Average.actual.speed..Akamai..kbps. ~ t$Price.for.very.high.speeds..combined)

lm(t$Average.actual.speed..Akamai..kbps. ~ t$Price.for.high.speeds..combined)

lm(t$Average.actual.speed..Akamai..kbps. ~ t$Price.for.med.speeds..combined)

lm(t$Average.actual.speed..Akamai..kbps. ~ t$Price.for.low.speeds..combined)
```

On average, the acctual speed for very high speed price is 7526.16.  The average speed for high speed price is 5339.31.
The average speed for medium speed is 6909.38. This speed is higher than the high speed internet.
The average speed for low speed price is 6178.61. Again, the average speed is higher than high speed internet.

Because of the problematic average speed for high price internet we added the percent of population in urban areas.

```{r}
lm(t$Average.actual.speed..Akamai..kbps. ~ t$Percent.of.population.in.urban.areas + t$Price.for.very.high.speeds..combined)
```

The average speed for very high speed internet price is 3788.05. For every dollar less the consumers paid, the speed reduced by 21.84.

```{r}
lm(t$Average.actual.speed..Akamai..kbps. ~ t$Percent.of.population.in.urban.areas + t$Price.for.high.speeds..combined)
```

The average speed for high speed internet is 3585.35. For every dollar less the consumers paid, the speed reduced by 21.84.

```{r}
lm(t$Average.actual.speed..Akamai..kbps. ~ t$Percent.of.population.in.urban.areas + t$Price.for.med.speeds..combined)
```

The average speed for medium speed internet is 4045.76. For every dollar less the consumers paid, the speed reduce by 79.98.

```{r}
lm(t$Average.actual.speed..Akamai..kbps. ~ t$Percent.of.population.in.urban.areas + t$Price.for.low.speeds..combined)
```
The average speed for low speed internet is 2987.39. For every dollar less the consumers paid for the internet, the average speed reduce by 83.42.


# Conclusion
In conclusion, the average speed is the highest for medium internet speed consumers. The result suggested that the price for high and very high speed did not have the advertised speed the internet providers suggested.
Other factors such as country, might also play a role in determining the average speed. 

